### 2023-01-04 elsman/manuscript.tex
updated response for R2 Comment 10

\mainmatter
\section{Introduction}
\label{sec:org497996e}
Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges. The International Zeolite Association (IZA) lists over 200 known zeolite frameworks that are described by dimensionality, pore shape and size, and Si/Al ratios \cite{baerlocher-database-nodate}. It is natural to characterize differences in zeolites by the size and shapes of the features present in the crystal. Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites (T-sites) of a zeolite. Rings have been used to identify feasible zeolites topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003,bermudez-calculation-2017}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints of physicochemical properties \cite{helfrecht-new-2019}. Rings can be used to describe the channel sizes for understanding shape selectivity in catalysis, zeolite building blocks for solid state chemists, and the sizes of framework cages and windows to provide insights into adsorption properties. The dimension and number of rings in a framework can be used to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites. The number and dimension  of rings that pass through a T-site can be used to describe and differentiate the local void environment around symmetry distinct T-sites and to  differentiate symmetry distinct oxygen atoms.

@@

A number of definitions and enumerations of rings in zeolites have been presented   \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001} and implemented in software tools, including ZeoTSites \cite{sastre-zeotsites-2001} and ToposPro \cite{blatov-applied-2014}. Zeolites are naturally  represented using graph theory, in which  atoms are nodes and bonds are edges.  Because the oxygen atoms are only connected to exactly two T-sites, their inclusion in a graph does not change the graph structure, and thus T-sites can be  defined as nodes and the intervening oxygen atoms as edges \cite{goetzke-properties-1991}.  The cutout of the chabazite (CHA) framework shown in \cref{fig:3.1}  and  \cref{tab:graphs} summarize basic definitions of graphs that will be used here.  The size of each of these features can either be defined by the total number of atoms contained in them, or by the number of T-sites contained in them. The latter method is the standard convention, which we will use. A path is a sequence of edges that connects a sequence of nodes in which every edge is distinct (shown with purple highlighted bonds in \cref{fig:3.1}). A cycle is a path that starts and ends at the same node, and does not repeat any other nodes (shown by blue bonds in \cref{fig:3.1}). A zeolite crystal contains an infinite number of cycles, but the symmetry of the crystal reduces this to a finite number of symmetry-distrinct rings whose exact identity is subject to definition. \cref{fig:3.1} shows an 8-membered ring (8-MR) and a 4-MR filled in pink and green, respectively.

>>>>

\mainmatter
\section{Introduction}
\label{sec:org497996e}
Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges. The International Zeolite Association (IZA) lists over 200 known zeolite frameworks that are described by dimensionality, pore shape and size, and Si/Al ratios \cite{baerlocher-database-nodate}. It is natural to characterize differences in zeolites by the size and shapes of the features present in the crystal. Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites (T-sites) of a zeolite. Rings have been used to identify feasible zeolites topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003,bermudez-calculation-2017}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints of physicochemical properties \cite{helfrecht-new-2019}. Rings can be used to describe the channel sizes for understanding shape selectivity in catalysis, zeolite building blocks for solid state chemists, and the sizes of framework cages and windows to provide insights into adsorption properties. The dimension and number of rings in a framework can be used to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites. Take for example the CHA framework, \cref{fig:3.1} shows an 8-membered ring (8-MR) and a 4-MR filled in with pink and green respectively. As an illustration of the significance of those rings, consider that the T-site labeled 20 is a Br\o nsted acid site, i.e., the site contains an Al and a charge-compensating proton bound to any one of the four adjacent oxygen atoms. While the T-site is symmetrically identical to all other T-sites in CHA, the proton sites are inequivalent, distinguished by the dimensions of the rings that pass through the oxygen atoms, a distinction that is experimentally observable in IR spectra \cite{kester-effects-2021}. The number and dimension  of rings that pass through a T-site can be used to describe and differentiate the local void environment around symmetry distinct T-sites and to  differentiate symmetry distinct oxygen atoms.

@@

A number of definitions and enumerations of rings in zeolites have been presented   \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001} and implemented in software tools, including ZeoTSites \cite{sastre-zeotsites-2001} and ToposPro \cite{blatov-applied-2014}. Zeolites are naturally  represented using graph theory, in which  atoms are nodes and bonds are edges.  Because the oxygen atoms are only connected to exactly two T-sites, their inclusion in a graph does not change the graph structure, and thus T-sites can be  defined as nodes and the intervening oxygen atoms as edges \cite{goetzke-properties-1991}.  The cutout of the chabazite (CHA) framework shown in \cref{fig:3.1}  and  \cref{tab:graphs} summarize basic definitions of graphs that will be used here.  The size of each of these features can either be defined by the total number of atoms contained in them, or by the number of T-sites contained in them. The latter method is the standard convention, which we will use. A path is a sequence of edges that connects a sequence of nodes in which every edge is distinct (shown with purple highlighted bonds in \cref{fig:3.1}). A cycle is a path that starts and ends at the same node, and does not repeat any other nodes (shown by blue bonds in \cref{fig:3.1}). A zeolite crystal contains an infinite number of cycles, but the symmetry of the crystal reduces this to a finite number of symmetry-distrinct rings whose exact identity is subject to definition. \cref{fig:3.1} shows an 8-MR and a 4-MR filled in pink and green, respectively.
